---
title: "World Maps"
author: 'Chris Bentz and Steven Moran'
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Session Info
Give the session info (reduced).
```{r, echo = F}
# R version
sessionInfo()$R.version$version.string
# platform
sessionInfo()$R.version$platform 
```

# Load Libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r, message = FALSE}
library(readr)
library(ggmap)
library(maps)
library(gridExtra)
library(ggrepel)
library(RCurl)
```

Give the package versions.
```{r, echo = F}
# version of packages loaded
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

# Load the language info 
Load language info file directly from the github repo.
```{r, message = F}
languages <- as.data.frame(read_csv("https://raw.githubusercontent.com/100LC/100LC/master/LangInfo/langInfo_100LC.csv?token=AAIGDLUYEVDIGWGOHXZK6QTBYG4V2"))
```
# All languages of 100 WALS sample

## Simple Stats
```{r}
length(unique(languages$iso639_3)) # number of languages according to iso
length(unique(languages$glottocode)) # number of languages according to glottolog
length(unique(languages$top_level_family)) # number of top level language families according to glottolog
length(unique(languages$family_wals)) # number of language families according to wals
unique(languages$macroarea_glotto) # number of macroareas according to glottolog (same as for WALS)
```
## Pre-Processing
```{r}
# add 360 to longitudes for languages with longitudes < -25 
# (this is necessary to create a Pacific centered map)
languages$longitude_wals[languages$longitude_wals < -25] <- 
  languages$longitude_wals[languages$longitude_wals < -25] + 360
# collapse status information into fewer factors
languages$status[languages$status %in% c("definitely endangered", "severely endangered", "critically endangered")]  <- "endangered" 
```

## World Map 
World maps with endangerment status information from Glottolog.
```{r, fig.width = 10, fig.height = 5, warning = FALSE}
# create world map
world <- map_data("world", wrap = c(-25, 335)) 
status.map <- ggplot() + 
  geom_polygon(data = world, aes(x = long, y = lat, group = group), 
               fill = "white", colour = "darkgrey") +
  geom_point(data = languages, aes(x = longitude_wals, y = latitude_wals, 
                                 fill = status), 
             alpha = 1, size = 3.5, pch = 21) +
             # select colours manually to be color blind safe
             scale_fill_manual(values = c("#E69F00", "#661100", "#009E73", "#F0E442")) +
  geom_text_repel(data = languages, aes(x = longitude_wals, y = latitude_wals, 
                                        label = name_wals), size = 2.5,
                  box.padding = unit(0.1, 'lines'), force = 0.5) +
  scale_y_continuous(limits = c(-65, 80)) +
  labs(x = "longitude", y = "latitude", fill = "Endangerment Status") +
  theme_bw() +
  theme(axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.position = "bottom")
status.map
```

Save to file.
```{r, warning = FALSE}
ggsave("~/Desktop/worldMap_100LC.pdf", status.map, 
       dpi = 300, scale = 1, width = 12, height = 6, device = cairo_pdf)
```
# Languages for which texts are currently available in the corpus

Exclude languages for which there is currently no data.
```{r}
# give selection of languages which are currently not represented by text data 
# (see folder Reports/line_counts/line_counts.csv on github)
missing <- c("Karok", "Koasati", "Koyraboro Senni", "Krongo", "Lakhota", "Lezgian", 
             "Mangarrayi", "Meithei", "Maricopa", "Slave", "Oneida", "Supiyre", "Tukang Besi")
# define a "no tin" operator
`%notin%` <- Negate(`%in%`)
# exclude these languages from the language info data frame
languages.corpus <- languages[languages$name_wals %notin% missing, ]
```

## Simple Stats
```{r}
length(unique(languages.corpus$iso639_3)) # number of languages according to iso
length(unique(languages.corpus$glottocode)) # number of languages according to glottolog
length(unique(languages.corpus$top_level_family)) # number of top level language families according to glottolog
length(unique(languages.corpus$family_wals)) # number of language families according to wals
unique(languages.corpus$macroarea_glotto) # number of macroareas according to glottolog (same as for WALS)
```

## Pre-Processing
```{r}
# add 360 to longitudes for languages with longitudes < -25 
# (this is necessary to create a Pacific centered map)
languages.corpus$longitude_wals[languages.corpus$longitude_wals < -25] <- 
  languages.corpus$longitude_wals[languages.corpus$longitude_wals < -25] + 360
```

## World Map 
World maps with family information.
```{r, fig.width = 10, fig.height = 7.5, warning = FALSE}
# create world map
world <- map_data("world", wrap = c(-25, 335)) 
family.map <- ggplot() + 
  geom_polygon(data = world, aes(x = long, y = lat, group = group), 
               fill = "white", colour = "darkgrey") +
  geom_point(data = languages.corpus, aes(x = longitude_wals, y = latitude_wals, 
                                 fill = family_wals), 
             alpha = 1, size = 3.5, pch = 21) +
             # select colors manually to be color blind safe
             # scale_fill_manual(values = c("#E69F00", "#661100", "#009E73", "#F0E442")) +
  geom_text_repel(data = languages.corpus, aes(x = longitude_wals, y = latitude_wals, 
                                        label = name_wals), size = 2.5,
                  box.padding = unit(0.1, 'lines'), force = 0.5) +
  scale_y_continuous(limits = c(-65, 80)) +
  labs(x = "longitude", y = "latitude", fill = "Family (WALS)") +
  theme_bw() +
  theme(axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.position = "bottom")
family.map
```

```{r, warning = FALSE}
ggsave("~/Desktop/worldMap_CurrentCorpus.pdf", family.map, 
       dpi = 300, scale = 1, width = 10, height = 7.5, device = cairo_pdf)
```