---
title: "Get the bibliographic references encoded in the TeDDi input files"
author: "Steven Moran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  pandoc_args: --webtex
---

Load the `clc` package R library and generate a table of references per file and per unique reference. 

Note this package is a work in progress. 

First load the TeDDi RData object.

```{r}
# TODO: we should put this online somewhere because I can't get it to load as part of the clc package
load('../../../clc/clc.RData')
```

```{r}
library(clc)
library(knitr)
library(readr)
```

Get the references from the TeDDi `file` database table.

```{r}
references <- clc::get_references()
```

Let's have a look at the results.

```{r}
references %>% head(n = 50) %>% kable()
```

Let's write these to a CSV file for easier inspection.

```{r}
write_csv(references, 'files-references.csv')
```

Now we get the unique language name and their bibliographic sources for corpora above, like Basque, which have lots of input text files from the same source, e.g. Open Subtitles.

```{r}
unique.references <- references %>% select(language_name_glotto, source) %>% distinct()
```

This leaves us with 1010 observations instead of a row for each file in the database.

```{r}
unique.references %>% head(n=20) %>% kable()
```

Let's write these to CSV as well.

```{r}
write_csv(unique.references, 'sources-references.csv')
```

Upon visual inspection, we see that some files have no source.

```{r}
unique.references %>% filter(source == "NA")
```

And others are quite limited in scope.

```{r}
unique.references %>% filter(language_name_glotto == "Wichita") %>% kable()
```

These should be udpated if they can, see:

* https://github.com/morphdiv/teddi_sample/issues/163

