---
title: "Get the bibliographic references encoded in the 100LC input files"
author: "Steven Moran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  pandoc_args: --webtex
---

Load the `clc` package R library and generate a table of references per file and per unique reference. 

Note this package is a work in progress. For example, it currently only loads the test version of the 100 LC database, i.e. `test.RData`. If you want to run the `clc::get_references()` function on the entire dataset, then you need to load the clc library (below) and then load the full database into RStudio.


```{r}
library(clc)
library(knitr)
```

Get the references from the 100LC `file` database table.

```{r}
references <- clc::get_references()
```

So let's write these to a CSV file for inspection.

```{r}
write_csv(references, 'files-references.csv')
```

Now we get the unique language name and their bibliographic sources for corpora above, like Basque, which have lots of input text files from the same source, e.g. Open Subtitles.

```{r}
unique.references <- references %>% select(language_name_glotto, source) %>% distinct()
```

This leaves us with 1010 observations instead of all files in the database.

Let's write these to CSV as well.

```{r}
write_csv(unique.references, 'sources-references.csv')
```
