---
title: "Get the unique mode and genre labels from metadata and file folder structure"
author: "Steven Moran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  pandoc_args: --webtex
---

# Setup  

```{r, message=FALSE}
library(dplyr)
library(readr)
library(knitr)
```

Load the RData serialized version of the TeDDi database.

```{r}
load('../../Database/TeDDi.RData')
```


# Get genre info from the files

Select the `mode`, `genre_narrow`, and `genre_broad` fields from the file table and get the distinct values per corpus.

```{r}
genre_info <- clc_file %>% select(corpus_id, language_name_wals, mode, genre_broad, genre_narrow) %>% distinct()
write_csv(genre_info, 'genres_in_file_metadata_headers.csv')
genre_info %>% kable()
```

What are the distinct values of each?

Mode:

```{r}
clc_file %>% select(mode) %>% distinct() %>% kable()
```

Genre broad:

```{r}
clc_file %>% select(genre_broad) %>% distinct() %>% kable()
```

Genre narrow:

```{r}
clc_file %>% select(genre_narrow) %>% distinct() %>% kable()
```

# Get genre info from the file folder structure

Select the `genre_narrow` and `genre_broad` categories as encoded in the file folder structure.

```{r}
file_folders <- clc_file %>% select(corpus_id, genre_broad, genre_narrow) %>% distinct()
corpus_names <- clc_corpus %>% select(id, name)
file_folders <- left_join(file_folders, corpus_names, by=c("corpus_id"="id"))
file_folders <- file_folders %>% select(name, genre_broad, genre_narrow)

write_csv(file_folders, 'genres_in_file_folders.csv')
kable(file_folders)
```

Get the unique values.

Genre broad:

```{r}
clc_file %>% select(genre_broad) %>% distinct() %>% kable()
```

Genre narrow:

```{r}
clc_file %>% select(genre_narrow) %>% distinct() %>% kable()
```
