---
title: "Generate tables for the 100LC database paper"
author: "Steven Moran\n"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  pandoc_args: --webtex
---

```{r, warning=FALSE, message=FALSE}
library(readr)
library(dplyr)
library(knitr)
library(xtable)
```

# Overview

Generate descriptive stats and plots for the 100LC database paper submission.


# Corpus overview

Load the 100LC index file.

```{r, warning=FALSE, message=FALSE}
index <- read_csv("../../LangInfo/langInfo_100LC.csv")
```

Let's add in the corpus sizes by word counts (as suggested [here](https://github.com/uzling/100LC/pull/193)).

```{r}
word_counts <- read_csv("../ttr/word_ttr.csv")
```

The output from the [word ttr script](https://github.com/uzling/100LC/blob/master/Reports/ttr/word_ttr.csv) is broken down by genre and writing systems when there are multiple scripts for the same corpus and genre, e.g. Vietnamese.

```{r}
word_counts %>%
  head(n = 15) %>%
  kable()
```

So let's first get the full counts.

```{r}
word_totals <- word_counts %>%
  group_by(name) %>%
  summarize(corpus_size = sum(tokens))
```

Next we can merge the counts into the 100LC index.

```{r}
index <- left_join(index, word_totals)
```

Decide what we want to keep from the language index and create a LaTeX (long) table for the paper submission.

```{r}
df <- index %>%
  select(name_glotto,
         glottocode,
         iso639_3,
         status,
         genus_wals,
         macroarea_glotto,
         corpus_size) %>%
  arrange(name_glotto)
```

Rename the columns.

```{r}
df <- df %>% rename("Language name" = name_glotto,
                    "Glottocode" = glottocode,
                    "ISO 630-3" = iso639_3,
                    "Endangerment" = status,
                    "Genus" = genus_wals,
                    "Area" = macroarea_glotto,
                    "Size (in words)" = corpus_size)
```

How about this for a descriptive table of the languages in the 100LC sample?


```{r}
df %>%
  head(n = 30) %>%
  kable()
```

Dump an xtable (`longtable` in latex) for the paper. Currently, I just copy and paste the table into the Overleaf document.

```{r}
add_to_row <- list(pos = list(0), command = NULL)
command <- paste0(
  "\\hline\n\\endhead\n",
  "\\hline\n",
  "\\multicolumn{", dim(df)[2], "}{l}",
  "{\\footnotesize Continued on next page}\n",
  "\\endfoot\n",
  "\\endlastfoot\n"
)
add_to_row$command <- command

print(xtable(df,
      digits = c(0, 0, 0, 0, 0, 0, 0, 0)),
      hline.after = c(-1),
      add_to_row = add_to_row,
      tabular.environment = "longtable",
      include.rownames = FALSE,
      file = "100LC_index.tex")

print(xtable(df,
      digits = c(0, 0, 0, 0, 0, 0, 0, 0)),
      hline.after = c(-1),
      add_to_row = add_to_row,
      tabular.environment = "longtable",
      include.rownames = FALSE)
```

# Word TTR

Now which stats for the paper (see: https://github.com/uzling/100LC/issues/189)?

Counts for both words and characters:

* number of types per genre and language
* number of tokens per genre and language

Calculations (per language and per genre and language):

* word TTR
* character TTR
* median word length (in characters)

Genres are represented by corpus IDs in the database (see: https://github.com/uzling/100LC/blob/master/Reports/genres/get_genres.md). This makes it easy to extract the pertinent file(s) per genre and to do the various type / token counts.

However, the 100LC corpus contains multiple writing scripts in files within the same language and genre (see: https://github.com/uzling/100LC/issues/189). This issue is discussed in this report on getting word types and tokens:

* https://github.com/uzling/100LC/blob/master/Reports/ttr/get_word_ttr.md

Let's load the results from that report.

```{r, warning=FALSE, message=FALSE}
word_ttr <- read_csv("../ttr/word_ttr.csv")
```

And have a look:

```{r}
word_ttr %>%
  head() %>%
  kable()
```

Let's rearrange the columns to make them more reader friendly.

```{r}
word_ttr <- word_ttr %>% select(name, genre_broad, writing_system, mean_word_length, types, tokens, ttr)
```

Rename the columns to match in the paper.

```{r}
word_ttr <- word_ttr %>%
  rename("Language" = name,
         "Genre" = genre_broad,
         "Writing system" = writing_system,
         "Word types" = types,
         "Word tokens" = tokens,
         "Mean word length" = mean_word_length,
         "Type-token ratio" = ttr)

word_ttr %>%
  head() %>%
  kable()
```

```{r}
add_to_row <- list(pos = list(0), command = NULL)
command <- paste0(
  "\\hline\n\\endhead\n",
  "\\hline\n",
  "\\multicolumn{", dim(word_ttr)[2] + 1, "}{l}",
  "{\\footnotesize Continued on next page}\n",
  "\\endfoot\n",
  "\\endlastfoot\n"
)
add_to_row$command <- command
print(xtable(word_ttr,
      digits = c(0, 0, 0, 0, 2, 0, 0, 2)),
      hline.after = c(-1),
      add_to_row = add_to_row,
      tabular.environment = "longtable",
      include.rownames = FALSE,
      file = "word_ttr.tex")

print(xtable(word_ttr,
      digits = c(0, 0, 0, 0, 2, 0, 0, 2)),
      hline.after = c(-1),
      add_to_row = add_to_row,
      tabular.environment = "longtable",
      include.rownames = FALSE)
```

# Grapheme TTR

We generate the grapheme TTR figures here:

* https://github.com/uzling/100LC/blob/master/Reports/ttr/get_grapheme_ttr.md

Let's load the results from that report.

```{r, warning=FALSE, message=FALSE}
grapheme_ttr <- read_csv("../ttr/grapheme_ttr.csv")
```

And have a look:

```{r}
grapheme_ttr %>%
  head() %>%
  kable()
```

Rename the columns to match in the paper.

```{r}
grapheme_ttr <- grapheme_ttr %>%
  rename("Language" = name,
         "Genre" = genre_broad,
         "Grapheme types" = types,
         "Grapheme tokens" = tokens,
         "Type-token ratio" = ttr)

grapheme_ttr %>%
  head() %>%
  kable()
```

```{r}
add_to_row <- list(pos = list(0), command = NULL)
command <- paste0(
  "\\hline\n\\endhead\n",
  "\\hline\n",
  "\\multicolumn{", dim(grapheme_ttr)[2], "}{l}",
  "{\\footnotesize Continued on next page}\n",
  "\\endfoot\n",
  "\\endlastfoot\n"
)
add_to_row$command <- command
print(xtable(grapheme_ttr,
             digits = c(0, 0, 0, 0, 0, 2)),
      hline.after = c(-1),
      add_to_row = add_to_row,
      tabular.environment = "longtable",
      include.rownames = FALSE,
      file = "grapheme_ttr.tex")
```

# Plots

For a [histrograph plot](https://github.com/uzling/100LC/pull/193) of word ttr, see [plot_ttr.md](../ttr/plot_ttr.md).

