---
title: "Graphemes exploration"
author: "Steven Moran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  pandoc_args: --webtex
---

```{r echo=FALSE, include=FALSE}
library(dplyr)
library(knitr)
```

This is the 100LC language info index file (planned languages, not actual languages in the corpus).

```{r}
index <- read.csv('../../LangInfo/langInfo_100LC.csv')
```

Which languages in 100LC are covered by Epitran, a tool for transliterating orthographic text as IPA:

* https://pypi.org/project/epitran/

Looks like 16 languages (but this depends on the script, which isn't indicated in the 100LC language info index):

```{r}
# This table was derived by hand from the project website above
epitran <- read.csv('epitran_data.csv')

table(index$iso639_3 %in% epitran$ISO6393) %>% kable()
```

```{r}
index[which(index$iso639_3 %in% epitran$ISO6393),] %>% select(iso639_3, name) %>% arrange(name) %>% kable()
```

Which languages have grapheme-to-phoneme coverage in The World Writing System Database (https://agricolamz.github.io/wwsd/)? What's the overlap between the two sources? (Note: we'll have to take into account whether the scripts are the same.)

```{r}
# wwsd tables
bib <- read.csv('~/Github/wwsd/bibliography.tsv', sep="\t", stringsAsFactors = FALSE)
db <- read.csv('~/Github/wwsd/database.csv', stringsAsFactors = FALSE)

table(index$glottocode %in% bib$glottocode) %>% kable()
```

```{r}
index[which(index$glottocode %in% bib$glottocode),] %>% select(iso639_3, name)  %>% arrange(name) %>% kable()
```