---
title: "Generate initial orthography profiles"
author: "Steven Moran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  pandoc_args: --webtex
---

# Overview

In its simpliest form, an orthography profile (Moran & Cysouw, 2018) is a Unicode character or Unicode grapheme unigram model, i.e. a CSV file where each row represents a character or grapheme in a writing system with additional information such as its frequency of occurrence in a text. This report generates orthography profiles from each corpus in the 100LC database. 

```{r echo=FALSE, include=FALSE}
# To install orthography profile tokenization:
# install.packages("devtools")
# devtools::install_github("cysouw/qlcData", build_vignettes = TRUE)

library(qlcData)
library(dplyr)
library(tidyr)
library(knitr)
library(purrrlyr)
```

I use the 100LC test data set for corpora for illustration purposes. The output for each corpus is written to the `orthography_profiles` folder in this directory. To run on the full database, load `100LC.Rdata`.

```{r}
# load('../../Database/test.RData')
load('../../Database/100LC.Rdata')
```

First, we need to merge in the corpus IDs with the lines table.

```{r}
lines <- clc_line %>% select(file_id, text)
corpus_ids <- clc_file %>% select(id, corpus_id) %>% distinct()
lines <- left_join(corpus_ids, lines, by=c("id"="file_id"))

# corpus_names <- tidyr::unite(clc_corpus, corpus, c(name, genre_broad))
# corpus_ids <- corpus_names %>% select(id, corpus) %>% distinct()
# lines <- left_join(corpus_ids, lines, by=c("id"="file_id"))
glimpse(lines)
```

Here are the frequency counts (word counts) per corpus. Recall that we need to do deal with languages that have multiple writing systems in the same folder. TODO.

```{r}
table(lines$corpus_id) %>% kable()
```

Here's how we generate an initial orthography profile (a linguistically correct orthography profile usually requires expert analysis of elements such as bigraphs; see Moran \& Cysouw, 2018 for detailed discussion).

```{r}
test <- lines %>% filter(id == 1)
op <- write.profile(test$text)
# Since we're working with lines, let's remove the spaces counted
op <- op %>% filter(Grapheme!=" ")
op$Frequency <- as.integer(op$Frequency)
op %>% arrange(desc(Frequency)) %>% kable()
```

```{r}
corpus_names <- tidyr::unite(clc_corpus, corpus, c(name, genre_broad))

x <- lines %>% 
  slice_rows("corpus_id") %>% 
  by_slice(function(x) as.vector(t(x[2])), .to = "words")

for (i in 1:nrow(x)) {
  row <- x[i,]
  op <- write.profile(row$words[[1]])
  op$Frequency <- as.integer(op$Frequency)
  op <- op %>% filter(Grapheme!=" ")
  op <- op %>% arrange(desc(Frequency))
  # Create the corpus name
  filename <- corpus_names %>% filter(id == i) %>% select(corpus)
  fname <- paste0("orthography_profiles/", filename$corpus, ".csv")
  write.csv(op, file=fname, row.names=FALSE)
}
```


# References

* Moran, Steven and Michael Cysouw. 2018. The Unicode Cookbook for Linguists: Managing writing systems using orthography profiles. Translation and Multilingual Natural Language Process- ing series in Language Science Press. DOI: https://doi.org/10.5281/zenodo.1296780. Online: http://langsci-press.org/catalog/book/176.