---
title: "Get language family counts for various multilingual resources"
author: "Steven Moran"
date: "(`r format(Sys.time(), '%d %B, %Y')`)"
output:
  github_document
---

```{r, message=FALSE}
library(tidyverse)
library(lingtypology)
library(testthat)
```

Get counts of families and area coverage for 100LC. This is easy because we have already combined the 100LC data with Glottolog.

Number of language families. As we discuss below, language family isolates have no `family_id` in Glottolog, and thus are listed `NA`. To account for this we need to add families to get the correct counts. In the case of 100LC, there are 9 language isolates in the current sample, so we add 8.

```{r, message=FALSE, warning=FALSE}
df <- read_csv('../../LangInfo/langInfo_100LC.csv')
nrow(df %>% filter(!is.na(family_id)) %>% select(family_id) %>% distinct()) + 8
```

Distribution of languages by area.

```{r}
table(df$macroarea_glotto)
```

For the other resources, we'll need the [Glottolog](https://glottolog.org) data.

```{r, message=FALSE, warning=FALSE}
glottolog <- read_csv('glottolog_languoid/languoid.csv')
glottolog_areas <- read_csv('https://cdstar.eva.mpg.de/bitstreams/EAEA0-E62D-ED67-FD05-0/languages_and_dialects_geo.csv')
```

Get [UD language](https://universaldependencies.org) family counts. Language codes are encoded in filenames. They can be two or three letter ISO 639 codes, so we'll need a mapping table to convert them all to 3-letter codes.

```{r, message=FALSE, warning=FALSE}
iso <- read_csv('language-codes-3b2_csv.csv')
```

Create the table.

```{r, message=FALSE, warning=FALSE}
ud <- read_csv('UD_fileprefixes.txt', col_names=F)
ud_split <- as.data.frame(str_split_fixed(ud$X1, "_", n=2))
ud <- cbind(ud, ud_split)
ud <- left_join(ud, iso, by=c("V1"="alpha2"))
ud <- ud %>% mutate(`alpha3-b` = coalesce(`alpha3-b`, V1))
ud <- ud %>% select(X1, V1, `alpha3-b`) %>% rename(file_prefix=X1, ISO6391 = V1, iso639P3code = `alpha3-b`)
ud <- left_join(ud, glottolog)
```

Some of the mappings from ISO639-1 to ISO639-3 are wrong. Note the NAs in the `family_id`. We need to fix these by hand.

```{r}
ud %>% select(file_prefix, ISO6391, iso639P3code, family_id) %>% head()
ud <- ud %>% select(file_prefix, ISO6391, iso639P3code, family_id)
write_csv(ud, 'ud.csv')
```

Relaod and remerge.

```{r}
ud <- read_csv('ud_by_hand.csv')
ud <- left_join(ud, glottolog)
ud <- left_join(ud, glottolog_areas, by=c("iso639P3code"="isocodes"))
```

We note that languagae isolates do not have a `family_id` in Glottolog, like Basque. This potentially throws our language family counts off, i.e. when there are two language isolates, counting them both as family `NA` would group them into the same "family". If thre is just one, then the counts are still correct (as in the case of UD).

```{r}
ud %>% filter(is.na(family_id))
```

How many language families are present?

```{r}
nrow(ud %>% filter(!is.na(family_id)) %>% select(family_id) %>% distinct())
```

Let's get the areas. UD is *heavily* Eurasian-centric.

```{r}
ud <- left_join(ud, glottolog_areas, by=c("iso639P3code"="isocodes"))
table(ud$macroarea)
```

Next, let's get the language family coverage for [mBERT](https://github.com/google-research/bert/blob/master/multilingual.md). We copy and paste the list of language names from [here](https://github.com/google-research/bert/blob/master/multilingual.md) and as it as [this csv file](mBert.csv).

```{r, message=FALSE}
mbert <- read_csv('mBERT.csv', col_names=F)
```

Now we can get some of the ISO 639-3 codes by language name look up. For this we can use the `lingtypology` library.

```{r, warning=FALSE, message=FALSE}
mbert_iso <- as.data.frame(lingtypology::iso.lang(mbert$X1))
mbert_iso <- mbert_iso %>% rownames_to_column(var="Name") %>% rename(iso639P3code = `lingtypology::iso.lang(mbert$X1)`)
```

Now we can get some of the ISO 639-3 codes by language name look up.

```{r, warning=FALSE, message=FALSE}
library(lingtypology)
mbert_iso <- as.data.frame(iso.lang(mbert$X1))
mbert_iso <- mbert_iso %>% rownames_to_column(var="Name") %>% rename(iso639P3code = `iso.lang(mbert$X1)`)
```

But doesn't get all of them, so we will have to fill in some by hand, which we do by first writing the data to disk and then manually adding the missing codes.

```{r}
write_csv(mbert_iso, 'mBERT_iso.csv')
```

Going through by hand, I randomly chose certain codes, e.g. Colloquiual Malay for "Malay", so that  they could be linked easily to Glottlog. Note also, that mBERT languages include things like:  Serbian, Bosnian, Croatian, Serbo-Croatian -- which by language code is simply "hbs". So the "language" coverage is actually less than what is reported.

Let's reload the by hand data.

```{r, message=FALSE}
mbert_iso <- read_csv('mBERT_iso_by_hand.csv')
expect_false(any(is.na(mbert_iso$iso639P3code)))
```

And calculate how many "true" languages the sample has.

```{r}
nrow(mbert_iso %>% select(iso639P3code) %>% distinct())
```

Now we merge in Glottolog data and calculate how many language families are present. Not many and less than UD. (Once again Basque is `NA` for `family_id`, but it's the only language isolate in the sample, so the counts are correct for language family.)

```{r}
mbert_iso <- left_join(mbert_iso, glottolog)
nrow(mbert_iso %>% select(family_id) %>% distinct())
```

How about the areas? mBERT languages are also heavily Eurasian-centric and it completely lacks any native South American languages.

```{r}
mbert_iso <- left_join(mbert_iso, glottolog_areas, by=c("iso639P3code"="isocodes"))
table(mbert_iso$macroarea)
```

By comparison here are the macroarea coverage in all of Glottolog.

```{r}
table(glottolog_areas$macroarea)
```

Now for the bibles data.

```{r}
bibles <- read_csv('bibles100codes.csv')
bibles <- left_join(bibles, glottolog)
bibles <- left_join(bibles, glottolog_areas)
```

How many languages according to ISO 639-3 codes?

```{r}
nrow(bibles %>% select(iso639P3code) %>% distinct())
```

How many language families? Since there are two islates in the sample, we need to add by one to account for the two `NA` in the `family_id`.

```{r}
nrow(bibles %>% select(family_id) %>% distinct()) + 1
```

Areal distribution? Note no data points in Australia.

```{r}
table(bibles$macroarea)
```

Now for XGLUE.

```{r}
xglue <- read_csv('xglue_languagecodes.txt', col_names = FALSE)
xglue <- xglue %>% rename(alpha2 = X1)
xglue <- left_join(xglue, iso)
xglue <- xglue %>% rename(iso639P3code = `alpha3-b`)
xglue <- left_join(xglue, glottolog)
```

Some of the XLGUE mappings are wrong. Note the NAs. We fix these by hand.

```{r}
xglue %>% select(family_id) %>% distinct()
tmp <- xglue %>% select(alpha2, iso639P3code, English, family_id)
# write_csv(tmp, 'xglue.csv')
tmp
```

Load the hand corrected XGLUE data.

```{r}
xglue <- read_csv('xglue_by_hand.csv')
```

Merge in Glottolog.

```{r}
xglue <- left_join(xglue, glottolog)
xglue <- left_join(xglue, glottolog_areas)
```

How many languages according to ISO 639-3 codes?

```{r}
nrow(xglue %>% select(iso639P3code) %>% distinct())
```

How many language families?

```{r}
expect_false(any(is.na(xglue %>% select(family_id) %>% distinct())))
nrow(xglue %>% select(family_id) %>% distinct())
```

Areal distribution? Note no data points in Australia. Most XGLUE languages are Eurasian. No Americas, Papunesia or Australia coverage.

```{r}
table(xglue$macroarea)
```
